---
title: "Regression of number of DE genes against cell counts"
author: "Ethan O'Farrell"
output:
  html_document:
  pdf_document:
---

```{r setup}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(ggplot2)
```

We suspect that a <cluster,drug,time> combination having more cell counts is associated
with finding more differentially expressed genes in that combination. Our reasoning
is that more cells will increase whatever confidence level we have in a give gene, 
no matter the actual differential expression value. 

Our solution to this is to randomly subsample cells such that each combination will
retain an equal number of cells. This analysis will tell us if that is necessary.

# Load Data

```{r}
df = read.csv('DE_cell_sensitivity.csv')
df$gene_fraction = df$gene / 24181
head(df)
```

# EDA

```{r, fig.width=8, fig,height=12}
p1 = df %>%
  ggplot(aes(x = cells, y = gene, color = name)) +
  geom_point()
p2 = df %>%
  ggplot(aes(x = cells, y = gene, color = de_direction)) +
  geom_point()

cowplot::plot_grid(p1, p2, ncol=1, align = "v")
```

We definitely need to consider DE Direction as a covariate

# Linear Model

$\mu = X\beta$ where $\mu$ is the number of genes designated differential

```{r}
m1 = lm(gene ~ cells + de_direction, data = df)
summary(m1)
par(mfrow = c(2,2))
plot(m1)
```

# Logistic Model

There is an upper and lower limit to the response variable. It can take on an integer
value between zero and the number of genes (24181). It would make sense to model 
this with logistic regression

$\text{logit} (\mu) = X\beta$ where $\mu$ is the fraction of genes designated differential

```{r}
m2 = glm(gene_fraction ~ cells + de_direction, data = df, family = binomial())
summary(m2)
par(mfrow = c(2,2))
plot(m2)
```
This model tells us the number of cells is NOT statistically significant for the
number of DE genes. This is unfortunate because even tho the other 2 models are 
both generally interpretable, this should theoretically be the most applicable
model.

# Poisson Model

Since the fraction of genes designated DE comes nowhere close to the upper limit,
and since this is integer data (not continuous), Poisson regression also is applicable.

$\log (\mu) = X\beta$ where $\mu$ is the number of genes designated differential

```{r}
m3 = glm(gene ~ cells + de_direction, data = df, family = poisson())
summary(m3)
par(mfrow = c(2,2))
plot(m3)
```

Residual diagnostics actually looks the best for the Poisson model, so we will adopt it.
The number of cells is a significant factor for the number of DE genes. This supports
our decision to normalize treatment combinations by cell count.